---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

# rsdv — The R Synthetic Data Vault

<!-- badges: start -->
[![R-CMD-check](https://github.com/kvenkita/rsdv/actions/workflows/check-standard.yaml/badge.svg)](https://github.com/kvenkita/rsdv/actions)
[![Codecov](https://codecov.io/gh/kvenkita/rsdv/branch/main/graph/badge.svg)](https://app.codecov.io/gh/kvenkita/rsdv)
<!-- badges: end -->

**Synthetic data generation in R (Gaussian Copula based, extensible to deep generative models)**

`rsdv` is an R implementation of Python's [Synthetic Data Vault (SDV)](https://sdv.dev/)
framework (Patki, Wedge, and Veeramachaneni 2016). It generates synthetic tabular data
using Gaussian copula models, with built-in quality and privacy evaluation.

## Installation

```r
# Development version
remotes::install_github("kvenkita/rsdv")
```

## Quick start

```{r example, warning=FALSE}
library(rsdv)

set.seed(42)

# Describe column types
meta <- metadata(adult_income) |>
  set_column_type("age",        "numerical") |>
  set_column_type("occupation", "categorical") |>
  set_column_type("income",     "categorical") |>
  set_primary_key("id")

# Fit a GaussianCopula synthesizer
syn       <- gaussian_copula_synthesizer(meta)
syn       <- fit(syn, adult_income)

# Generate 500 synthetic rows
synth_data <- sample(syn, n = 500)

# Evaluate quality
qr <- quality_report(real = adult_income, synthetic = synth_data,
                     metadata = meta)
print(qr)
```

`quality_report()` aggregates metrics into the two-property hierarchy used by
SDMetrics — **Column Shapes** (per-column marginal fidelity) and **Column Pair
Trends** (correlation similarity for numerical pairs, contingency similarity
for categorical pairs) — with the overall score the mean of the two.

`diagnostic_report()` complements it with structural-validity checks (value
ranges, category adherence, key uniqueness), and `sample_conditions()`
generates rows that hold given categorical values fixed:

```{r conditional, eval=FALSE}
# Validity checks
diagnostic_report(adult_income, synth_data, meta)

# Conditional generation
sample_conditions(syn, data.frame(income = ">50K", .n = 20))
```

## Related work

- Python SDV: [sdv-dev/SDV](https://github.com/sdv-dev/SDV)
- Synthetic Data Vault paper: Patki et al., IEEE DSAA 2016
- CTGAN: Xu et al., NeurIPS 2019 (implemented in companion package `rsdv.torch`)
